---
title: "AssessmentPart1"
author: "Mohammad Younes"
date: "10/30/2018"
output:
  pdf_document: default
  html_document: default
---

#QGIS Workflow - Generating a map of tree density in each London Ward

##1. Downloading Data:

1.1. Download OS Open Greenspace and OS Open Map Local ESRI Shape files from  OS Open Data. Select Tile TQ:

https://www.ordnancesurvey.co.uk/opendatadownload/products.html

1.2.	Download csv Street Tree Data from London Data Store: london_street_trees_gla_20180214.csv

https://data.london.gov.uk/dataset/local-authority-maintained-trees

1.3.	Download London Ward Data: London_Ward_CityMerged.shp

https://data.london.gov.uk/dataset/statistical-gis-boundary-files-london

##2.	Importing to Data QGIS:

2.1.	Open new QGIS project

2.2.	Add London Ward Layer: Layer > Add Layer > Add Vector Layer > Select London_Ward_CityMerged.shp

2.3.	Tree Data to Shapefile 
  2.4.1.	Import London_Tree CSV file: Layers > Add Layer > Add Delimited Text Layer. Select easting (X-field) and northing (Y-field)
  
  2.4.2.	Convert the CSV to a ShapeFile: Right click layer > Export > Save Selected layers as???  (make sure to select ESRI ShapeFile for file type)


2.4.3.	Import the London_Trees.shp into the project: Layers > Add Layer > Add Vector Layer
```{r}

setwd("Documents/TreeDensityRep/AssessmentPart1")
install.packages("png")
library(png)

img <- readPNG("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/Workflow Images/Image1.png")
print(img)

?readPNG
```

##3.	Data Analysis

3.1.	Go on QGIS and use the Count Points in Polygon Tool: Vector > Analysis Tool > Count Points in Polygon Tool.

3.1.1.	Select London_Ward_CityMerged for Polygons and London_Trees for Points.
3.1.2.	Give the new field a name i.e. TreeCount, then hit run
```{r}

img2 <- readPNG("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/Workflow Images/Image2.png)

```

3.1.3.	Save the output of the Points in Polygon tool as a Shapefile.

3.1.4.	Open Attribute table of the new layer and open field calculator.

3.1.6.	Create another new field calculating Tree per hectare

```{r}

img3 <- readPNG("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/Workflow Images/Image3.png")

```

3.1.7.	Save changes to the table.

##4 Set	Map Theme

4.1 Open properties of the Tree/Ha layer

4.2 Go to the symbology tab set color scheme to 'Graduated'
  4.2.1 Set column to Tree/Ha
  4.2.2 Set color ramp on Yellow to Green ('YlGn')
  4.2.3 Give the palette 9 classes
  
```{r}

img4 <- readPNG("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/Workflow Images/Image4.png")


```
  
##5 Produce Final Print out Map



## RStudio Work Flow

1. Import london_street_trees_gla_20180214.csv and convert to shp file, then set projection to British National Grid (ESPG 27700).

```{r}
library(sf)
library(rgdal)
library(tmaptools)
library(sp)

#Importing csv

London_Trees <- read.csv("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/London Data/london_street_trees_gla_20180214.csv", header = TRUE, sep = ",")

names(London_Trees)


#Setting coordinates to BNG
UKBNG <- "+init=epsg:27700"
London_Trees_Map <- st_as_sf(London_Trees, coords = c("easting", "northing"), crs = UKBNG)
class(London_Trees_Map)



#Exporting data frame to a shapefile
st_write(London_Trees_Map,"/Users/MedoYounes/Documents/UCL/CASA005/rMarkdown Work/London_Trees_Map.shp", driver = "ESRI Shapefile")

#Import the new shapefile
London_Trees_shp <- read_shape("/Users/MedoYounes/Documents/UCL/CASA005/rMarkdown Work/London_Trees_Map.shp")
names(London_Trees_shp)
London_Trees_shp <- st_as_sf(London_Trees_shp, coords = c("geometry"), crs = UKBNG)
set_projection(London_Trees_shp, UKBNG)
class(London_Trees_shp)
summary(London_Trees_shp)
qtm(London_Trees_shp, zoom = "high")
View(London_Tree_shp)

```

2. Import London_Ward_CityMerged.shp and set projection to British National Grid (ESPG 27700).

```{r}
library(tmap)
library(rgdal)

#read shp file from  local directory
London_Wards <-  read_shape("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/London Data/London-wards-2014 (1)/London-wards-2014_ESRI/London_Ward_CityMerged.shp")

#use UKBNG(previously defined) to provdie a CRS to the layer
set_projection(London_Wards,UKBNG)

qtm(London_Wards, fill = "white")


```

3. Count the number of trees in each ward using the poly_count tool, this will require the GISTools package.

```{r}
library(GISTools)
library(rgeos)
library(tidyverse)
library(tidyr)
library(tmap)
library(tmaptools)


Tree_Counts_Wards <- read_shape("/Users/MedoYounes/Documents/UCL/CASA005/GIS Coursework/London Data/London Trees/Ward_Tree_Count.shp")

data.frame(Tree_Counts_Wards)

                           

```

4. Plot the map using ggplot

```{r}
library(ggplot2)
library(RColorBrewer)


Tree_Counts_Wards <- fortify(Tree_Counts_Wards, region = "NAME")

#set the colour palette for the map, set low 'palegoldenrod' and high to 'green4'

palette1 <- scale_fill_continuous(low = "palegoldenrod", high = "green4", "Tree/Ha")

#set labels for map 
labels = labs(list(title = "Tree Density of London Wards", x = "Longitude", y = "Latitude"))

#plot the map using ggplot, making sure to include the pre-defined palette and labels
ggplot() + geom_sf(mapping = aes(geometry=geometry, fill = Tree.Ha), data = Tree_Counts_Wards) + theme_minimal() + palette1 + labels


```






```{r}


class(London_Trees_shp)

TreePoints <- data.frame(London_Trees$easting,London_Trees$northing)
TreePoints <- as.matrix(TreePoints)     
class(TreePoints)

Polygons(London_Wards)


SpatialPoints(TreePoints,proj4string = CRS(UKBNG))

Tree_Count <- poly.counts(TreePoints,London_Wards)


Tree_Counts <- as.data.frame(matrix(ncol = 9, nrow=length(London_Wards)))
names(London_Wards)
colnames(Tree_Counts) <- c("NAME", "GSS_CODE", "HECTARES", "NONLD_AREA", "LB_GSS_CD",  "BOROUGH", "POLY_ID", "geometry", "TreeCount")

Tree_Counts$NAME <- London_Trees_Map$NAME
Tree_Counts$GSS_CODE <- London_Trees_Map$GSS_CODE
Tree_Counts$HECTARES <- London_Trees_Map$HECTARES
Tree_Counts$NONLD_AREA <- London_Trees_Map$NONLD_AREA
Tree_Counts$LB_GSS_CD <- London_Trees_Map$LB_GSS_CD
Tree_Counts$BOROUGH <- London_Trees_Map$BOROUGH
Tree_Counts$POLY_ID <- London_Trees_Map$POLY_ID
Tree_Counts$geometry <- London_Trees_Map$geometry
Tree_Counts$GSS_CODE <- London_Trees_Map$BOROUGH

pt.poly <-over(London_Trees_shp,London_Wards)

```










